---
title: "Project part 2"
author: ""
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: yes
    code_folding: hide
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Load libraries that we use
```{r}
library(tidyverse)
library(data.table)
```

# Cleaning the data
Our goal is to predict if a machine part fails in general, therefore the names of the types of failure (TWF, HDF, etc.) becomes redundant, and we can aggregate all failure labels into a single column, `label`.  
```{r}
# import the data
df <- as.data.frame(read.csv("ai4i2020.csv"))
df <- df %>% select(-Product.ID) %>% 
  mutate(failure.type = 1)
head(df)

sum(is.na(df)) # Our data has no missing values
```

Here, we recode categorical values into numerical, scale the data, and drop redundant/unimportant column so we can focus on failure or no failure.  

The values for different features are on different scales, therefore we decided to scale the data.  
Note that, because the `Type` feature was originally categorical, we leave its numbers as is and only scale columns that contain measurements and would be meaningful to scale.

We are interested only if a machine part fails or not, therefore we drop specific types of failures and only use `Machine.failure`.
As an aside, it is possible for a machine part to fail due to more than one failure type. In our dataset, there were 24 of such machine parts.

Keep `UDI` so we can uniquely identify each observation.

```{r}
# Convert the string values for `Type` into numerical values: `L`, `M`, `H` into `1`, `2`, `3`, respectively.  
df$Type <- as.numeric(c('L'='1', 'M'='2', 'H'='3')[df$Type])
df[3:7] <- scale(df[3:7])

total_failures_per_part <- df %>% mutate(total = select(., 9:13) %>% rowSums()) %>% pull(total)
sum(total_failures_per_part>1) # machines that failed due to multiple fail types

# drop failure types
df <- df[1:8]

# clean column names
names(df) <- sub("[\\.]{2}.*", "", names(df))
head(df)

# export the cleaned data
#fwrite(df, "clean_data_scaled.csv")
```



